The title

### 2022-10-17 paper/manuscript.tex
update results
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

As stimuli, one piece of music, which was used in our previous experiment, was selected (for details, Tominaga et al., 2022). The piece was taken from Clementi's Sonatina Op.36 (No.3) in C major. The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable. The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only. Two expressive notations (i.e., articulation and dynamics) were added as \emph{Fig 1}.

@@

@@

First, we recorded participants' baseline performance by asking them to perform the piece without listening to any students' performance. A leading metronome (100 quarter beats per minute, 8 beats) was given before participants started performing the piece. Sheet music (\emph{Fig 1}) was displayed in front of the participants. Also, they were told to perform the piece expressively with their interpretation. This instruction was given to make sure that they paid attention to expressive aspects of the performance. Each participant performed the piece twice.

@@

We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present or absent). The \emph{aov\_ez} function in the \emph{afex} R package was used for a repeated-measures ANOVA. For post-doc comparisons on the estimated marginal means, we used the \emph{emmeans} R package.

@@

@@

>>>>

As stimuli, one piece of music, which was used in our previous experiment, was selected (for details, Tominaga et al., 2022). The piece was taken from Clementi's Sonatina Op.36 (No.3) in C major. The first 12 measures of the original piece were used and modified so that the piece had an almost equal number of data points for each dependent variable. The modified piece consisted of a 12-measure isochronous melody notated in 4/4 meter to be played with the right hand only. Two expressive notations (i.e., articulation and dynamics) were added as \emph{Fig \ref{fig:stim}}.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/stimli-1} \caption{\label{fig:stim}Sheet music. Articulation. For articulation notation, the curved line (slur) indicates legato and the dots indicate staccato. For dynamics notation, the symbol `f' denotes forte and the symbol `p' denotes piano. For data analysis, only the 8th notes were included.}\label{fig:stimli}
\end{figure}

@@

First, we recorded participants' baseline performance by asking them to perform the piece without listening to any students' performance. A leading metronome (100 quarter beats per minute, 8 beats) was given before participants started performing the piece. Sheet music (\emph{Fig \ref{fig:stim}}) was displayed in front of the participants. Also, they were told to perform the piece expressively with their interpretation and do their best as a performer. This instruction was given to make sure that they paid attention to expressive aspects of the performance. Each participant performed the piece twice.

@@

We performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Articulation (present vs.~absent) and Dynamics (present or absent) for each dependent variable (i.e.., IOIs, KOT, KV, KV Difference). The \emph{aov\_ez} function in the \emph{afex} R package was used for a repeated-measures ANOVA. For post-doc comparisons on the estimated marginal means, we used the \emph{emmeans} R package.

@@

All effects are reported as significant at \emph{p} \textless{} .05. For KOT, KV and KV Difference, we performed two-way ANOVAs separately for each subcomponent (i.e., legato, staccato, forte, piano).

\hypertarget{iois}{%
\subsection{IOIs}\label{iois}}

Neither main effect of Articulation (\emph{F}(1, 19) = 2.79, \emph{p} = 0.11, \(\eta_G^2\) = 0.010) and Dynamics (\emph{F}(1, 19) = 3.27, \emph{p} = 0.09, \(\eta_G^2\) = 0.003) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.48, \emph{p} = 0.24, \(\eta_G^2\) = 0.002). Therefore, participants did not change the tempo depending on the types of the recordings (\emph{Fig \ref{fig:ioi-1}}).

\hypertarget{kot}{%
\subsection{KOT}\label{kot}}

\hypertarget{legato}{%
\subsubsection{Legato}\label{legato}}

There was a significant main effect of Articulation (\emph{F}(1, 19) = 5.59, \emph{p} = 0.03, \(\eta_G^2\) = 0.002). However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.06, \emph{p} = 0.81, \(\eta_G^2\) = 0.000) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 1.89, \emph{p} = 0.19, \(\eta_G^2\) = 0.002). Participants produced longer legato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, left).

\hypertarget{staccato}{%
\subsubsection{Staccato}\label{staccato}}

There was a significant main effect of Articulation (\emph{F}(1, 19) = 4.88, \emph{p} = 0.04, \(\eta_G^2\) = 0.009). However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 0.68, \emph{p} = 0.42, \(\eta_G^2\) = 0.000) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 3.13, \emph{p} = 0.09, \(\eta_G^2\) = 0.001). Participants produced shorter staccato when listening to the recordings where articulation was not implemented (\emph{Fig \ref{fig:kot-1}}, right).

\hypertarget{kv}{%
\subsection{KV}\label{kv}}

\hypertarget{forte}{%
\subsubsection{Forte}\label{forte}}

Neither main effect of Articulation (\emph{F}(1, 19) = 1.66, \emph{p} = 0.21, \(\eta_G^2\) = 0.002) and Dynamics (\emph{F}(1, 19) = 0.33, \emph{p} = 0.57, \(\eta_G^2\) = 0.000) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.00, \emph{p} = 0.96, \(\eta_G^2\) = 0.000). Participants did not change performances in terms of dynamics (loudness) depending on the types of the recordings (\emph{Fig \ref{fig:vel-1}}, left).

\hypertarget{piano}{%
\subsubsection{Piano}\label{piano}}

There was a significant main effect of Articulation (\emph{F}(1, 19) = 7.18, \emph{p} = 0.01, \(\eta_G^2\) = 0.060). However, there was no significant main effect of Dynamics (\emph{F}(1, 19) = 1.80, \emph{p} = 0.20, \(\eta_G^2\) = 0.003) or interaction between Articulation and Dynamics (\emph{F}(1, 19) = 0.04, \emph{p} = 0.85, \(\eta_G^2\) = 0.000). Participants produced softer piano when listening to the recordings where articulation was implemented (\emph{Fig \ref{fig:vel-1}}, right).

\hypertarget{kv-difference}{%
\subsection{KV Difference}\label{kv-difference}}

\hypertarget{forte-to-piano}{%
\subsubsection{Forte to Piano}\label{forte-to-piano}}

Neither main effect of Articulation (\emph{F}(1, 19) = 3.40, \emph{p} = 0.08, \(\eta_G^2\) = 0.012) and Dynamics (\emph{F}(1, 19) = 1.58, \emph{p} = 0.22, \(\eta_G^2\) = 0.002) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 0.97, \emph{p} = 0.34, \(\eta_G^2\) = 0.001).

\hypertarget{piano-to-forte}{%
\subsubsection{Piano to Forte}\label{piano-to-forte}}

Neither main effect of Articulation (\emph{F}(1, 19) = 2.25, \emph{p} = 0.15, \(\eta_G^2\) = 0.005) and Dynamics (\emph{F}(1, 19) = 3.76, \emph{p} = 0.07, \(\eta_G^2\) = 0.004) nor the interaction between Articulation and Dynamics was significant (\emph{F}(1, 19) = 1.08, \emph{p} = 0.31, \(\eta_G^2\) = 0.001).

These results indicated that participants did not change performances in terms of dynamics contrast depending on the types of the recording (\emph{Fig \ref{fig:vel-diff-1}}).

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-1-1} \caption{\label{fig:ioi-1}IOIs (ms). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-1}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-1-1} \caption{\label{fig:kot-1}KOT(ms) for each subcomponent; legato (left) and staccato (right). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-kot-1}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-1-1} \caption{\label{fig:vel-1}KV (0-127) for each subcomponent; forte (left) and piano (right). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-1}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-1-1} \caption{\label{fig:vel-diff-1}KV Difference (-127-127) for each subcomponent; forte to piano (left) and piano to forte (right). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-diff-1}
\end{figure}

\hypertarget{comparison-with-baseline-performance}{%
\subsection{Comparison with baseline performance}\label{comparison-with-baseline-performance}}

In order to investigate how their performances during the experiment differed from their baseline performances (i.e., performances that participants produced before listening to any of the recordings), we performed one-way repeated-measures ANOVAs. For KOT, KV and KV Difference, we performed ANOVAs separately for each subcomponent. To recall the instruction for the baseline performance, we asked participants to perform the piece expressively with their interpretation and do their best as a performer. By comparing with the baseline performance, we can investigate how didactic intentions influence participants' performances.

\hypertarget{iois-1}{%
\subsection{IOIs}\label{iois-1}}

We compared the baseline performance with the performances where either both articulation and dynamics were implemented or neither of them was implemented to examine how participants performed differently for these extreme examples of students' skill levels. We categorised performances into three groups (baseline, both, none) and treated them as a factor Category.

There was no significant main effect of Category (\emph{F}(1.11, 21.18) = 0.58, \emph{p} = 0.47, \(\eta_G^2\) = 0.012; Greenhouse--Geisser corrected), suggesting that participants kept the tempo regardless of whether they had an intention to teach (\emph{Fig \ref{fig:ioi-2}}).

\hypertarget{kot-1}{%
\subsection{KOT}\label{kot-1}}

We compared the baseline performance with the performances where dynamics was implemented to examine how participants performed differently depending on whether articulation was implemented or not on each recording. We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), dynamics-only (i.e., articulation-absent, dynamics-present)) and treated them as a factor Category.

\hypertarget{legato-1}{%
\subsubsection{Legato}\label{legato-1}}

There was a significant main effect of Category (\emph{F}(1.45, 27.53) = 11.82, \emph{p} = 0.00, \(\eta_G^2\) = 0.060; Greenhouse--Geisser corrected). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} \textless{} .001 , baseline and dynamics-only: \emph{p} = .008), suggesting that participants produced longer legato when they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, left).

\hypertarget{staccato-1}{%
\subsubsection{Staccato}\label{staccato-1}}

There was no significant main effect of Category (\emph{F}(1.12, 21.28) = 0.75, \emph{p} = 0.41, \(\eta_G^2\) = 0.011; Greenhouse--Geisser corrected), suggesting that participants did not play staccato differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:kot-2}}, right).

\hypertarget{kv-1}{%
\subsection{KV}\label{kv-1}}

We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not on each recording. We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.

\hypertarget{forte-1}{%
\subsubsection{Forte}\label{forte-1}}

There was no significant main effect of Category (\emph{F}(1.59, 30.18) = 1.43, \emph{p} = 0.25, \(\eta_G^2\) = 0.004; Greenhouse--Geisser corrected), suggesting that participants did not play forte differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, left).

\hypertarget{piano-1}{%
\subsubsection{Piano}\label{piano-1}}

There was a significant main effect of Category (\emph{F}(1.79, 34.09) = 5.68, \emph{p} = 0.01, \(\eta_G^2\) = 0.072; Greenhouse--Geisser corrected). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there were differences between baseline and the other two categories (baseline and both: \emph{p} = .05 , baseline and articulation-only: \emph{p} = .03), suggesting that participants produced softer piano when they had an intention to teach (\emph{Fig \ref{fig:vel-2}}, right).

\hypertarget{kv-difference-1}{%
\subsection{KV Difference}\label{kv-difference-1}}

We compared the baseline performance with the performances where articulation was implemented to examine how participants performed differently depending on whether dynamics was implemented or not on each recording. We categorised performances into three groups (baseline, both (i.e., articulation-present, dynamics-present), articulation-only (i.e., articulation-present, dynamics-absent)) and treated them as a factor Category.

\hypertarget{forte-to-piano-1}{%
\subsubsection{Forte to Piano}\label{forte-to-piano-1}}

There was no significant main effect of Category (\emph{F}(1.53, 29.15) = 1.73, \emph{p} = 0.20, \(\eta_G^2\) = 0.014; Greenhouse--Geisser corrected).

\hypertarget{piano-to-forte-1}{%
\subsubsection{Piano to Forte}\label{piano-to-forte-1}}

There was no significant main effect of Category (\emph{F}(1.28, 24.39) = 2.27, \emph{p} = 0.14, \(\eta_G^2\) = 0.013; Greenhouse--Geisser corrected).

These results indicated that participants did not make dynamics contrast between forte and piano differently depending on whether they had an intention to teach (\emph{Fig \ref{fig:vel-diff-2}}).

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-2-1} \caption{\label{fig:ioi-2}Comaprison with the baseline performance in terms of IOIs (ms). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-2}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-kot-2-1} \caption{\label{fig:kot-2}Comaprison with the baseline performance in terms of KOT(ms) for each subcomponent; legato (left) and staccato (right). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-kot-2}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-2-1} \caption{\label{fig:vel-2}Comaprison with the baseline performance in terms of KV (0-127) for each subcomponent; forte (left) and piano (right). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-2}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-2-1} \caption{\label{fig:vel-diff-2}Comaprison with the baseline performance in terms of KV Difference (-127-127) for each subcomponent; forte to piano (left) and piano to forte (right). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-vel-diff-2}
\end{figure}

@@

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{acknowledgement}{%
\section{Acknowledgement}\label{acknowledgement}}
